LME-MPM applied to quasi-brittle fracture.

### 2020-05-19 _region_.tex
Cambios 19-5-2020

\message{ !name(2020_EFM_MPM_Eigensoftening.tex)}\documentclass[preprint,12pt,a4paper]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}

\usepackage{algcompatible}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{algpseudocode}

\usepackage{cancel}

@@

as follows

@@

material point $\vec{x}_p^{k+1}$:

@@

where $m_p$ can be computed as:

@@

This definition of the equivalent critical stress allow us to
link it with the averaged strain energy as $\delta W_{\epsilon,p} =
 \tens{\sigma}_{\epsilon,p}\ \delta\varepsilon_p$. The fracture
criterion is defined as once $\tens{\sigma}_{\epsilon,p}^{k+1}$
surpasses the tensile strength, $f_t$, the softening behaviour is
activated, which in turn causes a reduction of the initial forces as,

@@

   f^{int}_I = \sum_p (1 - \chi_p)\ \tens{\sigma}_{p}^{k+1} \cdot \Grad{N_{Ip}}

@@

where $\chi_p$ is the damage variable of each material point $p$,
ranges between zero (an intact material) and one (completely failed
material points). For the case of a linear softening such the sketched
in the Figure \ref{fig:Damage-ft-wc}, it is calculated as,

@@

Following Bazant~\cite{Bazant83} crack-model, \cite{Navas_2017_ES}
introduced a band width $h_{\epsilon}$, which has values between two
and four times the maximum size of the aggregates for the
concrete. The effective fracture strain $\varepsilon_{\epsilon,f}$ is
defined as the difference between the strain at crack initialization,
$\varepsilon_1(\vec{x}_p^{0})$, and the current strain, $\varepsilon_1(\vec{x}_p^{k+1})$, for material
point $p$. Also, $\varepsilon_{\epsilon,f}$ can be represented as the
current crack opening $w_n$ within the band width,
$h_{\epsilon}$. Therefore,
\begin{equation}
  \label{eq:effective-fracture-strain}
  \varepsilon_{\epsilon,f} = \varepsilon_1(\vec{x}_p^{k+1}) -
  \varepsilon_1(\vec{x}_p^{0}) = \frac{w_n}{h_{\epsilon}}
\end{equation}
Introducing \eqref{eq:effective-fracture-strain} in
\eqref{eq:damaged-variable-chi}, the damage variable can be computed
as,
\begin{equation}
  \label{eq:damage-variable-chi-II}
\chi = \frac{\varepsilon_{\epsilon,f}\ h^{\epsilon}}{w_c}  
\end{equation}
for the case of a linear softening behaviour. In a more general case,
the damage variable can be expressed in terms of the following
variables,
\begin{equation}
  \label{eq:damage-variable-chi-III}
  \chi = \chi(\varepsilon_{\epsilon,f}, h^{\epsilon}, f_t, w_c, G_f)
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

@@

This problem is devoted to compare the develop of both linear shape
functions and the \acrshort{lme}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figures/Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figures/Reactions_displacements_square_panel_mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:6}

@@

\begin{algorithm}
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
  \begin{algorithmic}
    \Require Particle status\\
    Number of particles: $N_p$\\
    $\epsilon$-neighbourhood of each particle $p$ : $B_{\epsilon,p}$\\
    \Require Material data\\
    Tensile strength: $f_{t,p}$\\
    Bandwidth of the cohesive fracture: $h_{\epsilon,p}$ \\
    Critical opening displacement: $w_c$\\ 
    \Ensure Return damage parameter $\chi := \{\chi_p\}$
    \State $\chi_p \leftarrow \chi_p^{k}$
    \For{$p$ to $N_p$}
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    \For{$q \in B_{\epsilon,p}$}
    \If{$\chi_q < 1$}    
    \State $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$
    \EndIf    
    \State $m_p \leftarrow m_p + m_q$
    \EndFor
    \State $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$
    \If{$\sigma_{p,\epsilon} > f_{t,p}$}
    \State $\epsilon_{f,p} = \epsilon_{I,p}$   
    \EndIf        
    \Else[$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$]
    \State $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
    \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c} \} \Big \}$    
    \EndIf    
    \EndFor
  \end{algorithmic}
\end{algorithm}

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-813) }

>>>>

\message{ !name(2020_EFM_MPM_Eigensoftening.tex)}% \documentclass[preprint,12pt,a4paper]{elsarticle}
\documentclass[12pt,a4paper]{elsarticle}
\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{subfigure} % subfiguras
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

 to

@@

material point $\vec{x}_p^{k+1}$as

@@

where $m_p$ is the total mass of the $\epsilon$-neighborhood

@@

This definition of the equivalent critical stress leads to a
definition of the averaged strain energy in terms of the averaged
strain as $\delta W_{\epsilon,p} =
 \tens{\sigma}_{\epsilon,p}\ \delta\varepsilon_p$. The softening behaviour is
activated once $\tens{\sigma}_{\epsilon,p}^{k+1}$ surpasses the
tensile strength, $f_t$. This consists in a reduction of the internal
forces as,

@@

   f^{int}_I = \sum_p (1 - \chi_p)\ \tens{\sigma}_{p}^{k+1} \cdot
   \Grad{N_{Ip}}\ \Omega_p

@@

where $\chi_p$ and $\Omega_p$ are respectively the damage or softening
variable and the volume for each material point $p$. $\chi_p$ takes
values between zero (an intact material) and one (completely failed
material points). For the case of a linear softening such the sketched in the Figure \ref{fig:Damage-ft-wc}, $\chi_p$ is computed as,

@@

In analogy to the band crack model presented by
Bazant~\cite{Bazant83}, Navas {\it et al.} \cite{Navas_2018_ES}
\cite{Navas_2017_ES} introduced a band width parameter $h_{\epsilon}$.
Concerning this parameter, a typical value of it is between two and
four times the maximum size of the aggregates, in the case of
concrete as brittle material. The effective fracture strain
$\varepsilon_{\epsilon,f}$ is defined as the difference between the strain at crack initialization,
$\varepsilon_1(\vec{x}_p^{0})$, and the current strain, $\varepsilon_1(\vec{x}_p^{k+1})$, for material point $p$. Also,
$\varepsilon_{\epsilon,f}$ can be represented as the current crack
opening $w_n$ within the band width, $h_{\epsilon}$. Therefore, 
  \label{eq:effective-fracture-strain}
  \varepsilon_{\epsilon,f} = \varepsilon_1(\vec{x}_p^{k+1}) -
  \varepsilon_1(\vec{x}_p^{0}) = \frac{w_n}{h_{\epsilon}}
Introducing \eqref{eq:effective-fracture-strain} in
\eqref{eq:damaged-variable-chi}, the damage variable can be computed
as,
  \label{eq:damage-variable-chi-II}
\chi = \frac{\varepsilon_{\epsilon,f}\ h^{\epsilon}}{w_c}.  
The function of $\chi$ presented in \eqref{eq:damage-variable-chi-III}
represents a linear softening behaviour. For a general case, the
damage variable can be expressed in terms of the following variables,
  \label{eq:damage-variable-chi-III}
  \chi = \chi(\varepsilon_{\epsilon,f}, h^{\epsilon}, f_t, w_c, G_f)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings. The first is the presence of stress instabilities even
with \acrshort{gimp} shape functions. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
impressive performance mitigating spurious stress oscillations
\cite{Wobbes2020} under the \acrshort{mpm} framework help to enhance
notoriously the quality of the results as we can see in Section
\ref{sec:edge-cracked-square-I}. Where both interpolation techniques
are compared by carry out an eigenerosion simulation. The second
limitation is concerning its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved thorough the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Section \ref{sec:3.3}, where experimental results
of a drop-weight impact test are compared with those produced by
eigensoftening computations.

@@

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figures/Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus $E = 1.06$, the Poisson's ratio
$\mu = 0.333$, and critical energy-release rate $G_c = 0.0001$.

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the loading and also in the unloading regions
of the curve when a linear interpolation technique is employed. In
contrast with it \acrshort{lme} simulation does not exhibit this
accuracy and remains rectilinear during loading and unloading. An
important consideration regarding the presence of oscillations once
both parts of the panel are separated. Is that they should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the simulation.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figures/Reactions_displacements_square_panel_mode_I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}

\begin{figure}
\centering
\subfigure[t1]{\includegraphics[width=0.7\textwidth]{./Figures/Edge_cracked_square_panel_in_mode_I_Stress_0100.png}}
\subfigure[t2]{\includegraphics[width=0.7\textwidth]{./Figures/Edge_cracked_square_panel_in_mode_I_Stress_0250.png}}
\subfigure[t3]{\includegraphics[width=0.7\textwidth]{./Figures/Edge_cracked_square_panel_in_mode_I_Stress_0398.png}}
\caption{Legos.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}

Edge_cracked_square_panel_in_mode_I_Stress_0250.png
Edge_cracked_square_panel_in_mode_I_Stress_0398.png

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:4}

@@

\begin{algorithm}
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
\end{algorithm}

@@

\end{document}
\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-867) }
